Reconstruction of Large Scale Structure by Nonlinear Displacement Fields

### 2016-09-24 ms_disp1/ms_disp1.tex
haoran

\begin{abstract}
We investigate the ability of using nonlinear displacement field to reconstruct the primordial linear perturbations of the large scale structure. According to the linear Lagrangian perturbation theory, the first order primordial linear density field is given by the negative divergence of the displacement field. By running $N$-body simulations, we show that this reconstruction algorithm recovers great amount of nonlinear information in the early universe. In reality, the displacement fields can be solved from only the final stage of the density fields. This has potential to reconstruct baryonic acoustic oscillation (BAO) from current and future large scale structure surveys.
\end{abstract}

@@

These algorithms include nonlinear Wiener filters in wavelet space, distribution function transforms etc. Standard BAO reconstruction algorithms smooths the nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large scale bulk flows by a negative Zel'dovich linear displacement. Here we propose a new reconstruction method that uses the nonlinear displacement field to recover the primordial density field. In the linear Lagrangian perturbation theory, the negative divergence of the the displacement field respect to Lagrangian coordinates gives the linear density field. Of course, the full displacement field is non-observable, as it requires the initial distributions of matter, however there are many techniques to estimate the nonlinear displacement field from a final distribution of matter. For example, when a homogeneous initial matter distribution is assumed, there is a unique solution of curl-less displacement field to relate the initial and final distributions without shell-crossing. This solution can be solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}. In 1-dimensional case, the solution simplifies to a reordering of matter elements by Eulerian positions, and [Hong-ming et. al.] studies this case. The homogeneous and curl-less assumptions are generally valid in cosmology [cite or proof]. Shell-crossing does exist, and this effect can be quantified by $N$-body simulations.

In this paper we track the initial and final distributions of particles in LSS $N$-body simulations, and use the displacement field to reconstruct the initial linear density field. The rest of this paper is structured as follows. In section \ref{sec.method} we describe the simulation and reconstruction algorithm. In section \ref{sec.results} we show the results of the reconstruction. Discussion and conclusion are in section \ref{sec.discussion}.

\begin{figure}[b] \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
  \label{fig.1}
\end{figure}
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig2.pdf}
  \caption{Caption goes here.}
  \label{fig.2}
\end{figure}

@@

\subsection{Simulation}
We use the public cosmological simulation code {\tt CUBE}. Cosmological parameters are in accordance with Planck. Initial conditions are generated at redshift $z=50$ using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates ${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ to describe the density fluctuations. The primordial linear density field $\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth factor.

Two-point statistics of the density fields are quantified by the cross power spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)\delta_j(k)|\rangle$, where subscripts $i,j$ may refer to linear, nonlinear, or reconstructed density fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$. The blue solid and black dashed curves in Fig.\ref{fig.1} show the power spectra of $\delta$ and $\delta_L$. Their difference shows the nonlinear evolution of LSS on small scales. Their cross power drops to a very low value, indicating a loss of linear information in the nonlinear power spectrum.

\subsection{Reconstruction}
In the simulation, we uses particle-ID (PID) to record the initial location ${\bs q}$ of particles, and the information is tracked until the $z=0$ and we can get the Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every particle. Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).

To quantify the linear information in the reconstructed density field, we decompose $\delta_R$ in Fourier space as
\begin{equation}\label{eq.decompose}
    \delta_R(k)=r'\delta_L+\delta_N,
\end{equation}
where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_L$ gives
\begin{equation}
    P_{LR}=r'P_{LL}+P_{LN},
\end{equation}
where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power spectrum. Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$. With the definition of cross correlation coefficient $r\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$. We plot the cross correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.2}. Clearly, $\delta_R$ contains much more linear information on smaller scales.

@@

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig3.pdf}
  \caption{Caption goes here.}
  \label{fig.3}
\end{figure}

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig4.pdf}
  \caption{Caption goes here.}
  \label{fig.4}
\end{figure}

\section{Results}\label{sec.results}
Fig.\ref{fig.2} shows the damping factors $W^2(k)$ for the optimal filtered nonlinear and reconstructed density fields. We fit the Gaussian BAO damping model ${\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ Mpc$/h$ for nonlinear and reconstructed fields.

>>>>

\begin{abstract}
We investigate the ability of using nonlinear displacement field to reconstruct 
the primordial linear perturbations of the large scale structure. According to 
the linear Lagrangian perturbation theory, the first order primordial linear 
density field is given by the negative divergence of the displacement field. By 
running $N$-body simulations, we show that this reconstruction algorithm recovers 
great amount of nonlinear information in the early universe. In reality, the 
displacement fields can be solved from only the final stage of the density 
fields. This has potential to reconstruct baryonic acoustic oscillation (BAO) 
from current and future large scale structure surveys.

\end{abstract}

@@

These algorithms include nonlinear Wiener filters in wavelet space, distribution 
function transforms etc. Standard BAO reconstruction algorithms smooths the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement. Here we propose a 
new reconstruction method that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field respect to Lagrangian coordinates 
gives the linear density field. Of course, the full displacement field is non-
observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..
115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
matter elements by Eulerian positions, and \cite{2016arXiv160907041Z} studies this 
case. The homogeneous and curl-less assumptions are generally valid in cosmology 
[cite or proof]. Shell-crossing does exist, and this effect can be quantified by 
$N$-body simulations.

In this paper we track the initial and final distributions of particles in LSS $N$-
body simulations, and use the displacement field to reconstruct the initial linear 
density field. The rest of this paper is structured as follows. In section 
\ref{sec.method} we describe the simulation and reconstruction algorithm. In 
section \ref{sec.results} we show the results of the reconstruction. Discussion and 
conclusion are in section \ref{sec.discussion}.

\begin{figure}[b] \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
  \label{fig.1}
\end{figure}
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig2.pdf}
  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$.}
  \label{fig.2}
\end{figure}

@@

\subsection{Simulation}
We use the public cosmological simulation code {\tt CUBE}. Cosmological parameters 
are in accordance with Planck. Initial conditions are generated at redshift $z=50$ 
using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via 
their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ 
Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
\cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.
436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force 
calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates 
${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ 
to describe the density fluctuations. The primordial linear density field $
\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth 
factor.

Two-point statistics of the density fields are quantified by the cross power 
spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)||\delta_j(k)|\rangle$, 
where subscripts $i,j$ may refer to linear, nonlinear, or reconstructed density 
fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We 
usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$. 
The blue solid and black dashed curves in Fig.\ref{fig.1} show the power spectra of 
$\delta$ and $\delta_L$. Their difference shows the nonlinear evolution of LSS on 
small scales. Their cross power drops to a very low value, indicating a loss of 
linear information in the nonlinear power spectrum.

\subsection{Reconstruction}
In the simulation, we uses particle-ID (PID) to record the initial location ${\bs 
q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
particle. Then these vectors are interpolated onto the initial Lagrangian 
coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter 
elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).

To quantify the linear information in the reconstructed density field, we decompose $\delta_R$ in Fourier space as
\begin{equation}\label{eq.decompose}
    \delta_R(k)=r'\delta_L+\delta_N,
\end{equation}
where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_L$ gives
\begin{equation}
    P_{LR}=r'P_{LL}+P_{LN},
\end{equation}
where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power 
spectrum. Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$. With the 
definition of cross correlation coefficient $r(\delta_L,\delta_R)\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ 
and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$. We plot the cross 
correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.2}. Clearly, $
\delta_R$ contains much more linear information on smaller scales.

@@

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig3.pdf}
  \caption{Power spectra of the nonlinear density field and its
  filtered density field.}
  \label{fig.3}
\end{figure}

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig4.pdf}
  \caption{Power spectra of the reconstructed density field and its
  filtered density field.}
  \label{fig.4}
\end{figure}

\section{Results}\label{sec.results}
Fig.\ref{fig.2} shows the cross correlation functions,  damping factors $W^2(k)$ for the optimal filtered 
nonlinear and reconstructed density fields. We fit the Gaussian BAO damping model $
{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
Mpc$/h$ for nonlinear and reconstructed fields.
